---
title: "Software For Scientists (1st Day Afternoon)"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Greetings from Maine

## Functions

### What is a function?

Functions gather a sequence of operations into a whole, preserving it for ongoing use. Functions provide:

* a name we can remember and invoke it by
* relief from the need to remember the individual operations
* a defined set of inputs and expected outputs
* rich connections to the larger programming environment

Let us start with a function adding two numbers:

```{r addition}
# addition of two numbers
add <- function(a, b) {
  return(a+b)
}
```

#### Exercise 1: 
Write a function of two arguments *a* and *b* that returns *a-b*, for example

```{r substraction, echo=FALSE}
# substract of two numbers
substract <- function(a, b) {
  return(a-b)
}
```

```{r example1}
substract(a=10, b=6)
```

#### Exercise 1 extras:

First, try to predict what happens. Then copy the code to R to see the answer:

```{r example1p, eval=FALSE}
substract(10, 6)
substract(b=6, a=10)
substract(b=6, 10)
b<-6; substract(a=10)
```


## Diversity Outbred Phenotypes

 <http://churchill.jax.org/research/cc/doresources.shtml>.


```{r dataset}
# bit.ly/mousepheno is a shortcut for svenson_149.csv dataset
# see https://github.com/simecek/dc_2016_10 
dt <- read.csv("http://bit.ly/mousepheno")
```

## Acknowledgements

